Formal Grammars as Types in Non-commutative Bunched Type Theory

### 2023-12-29 paper/paper.tex
added comment re:glued model

\newcommand{\cat}[1]{\mathbf{#1}}
\newcommand{\Set}{\mathbf{Set}}
\newcommand{\Type}{\mathbf{Type}}
\newcommand{\Prop}{\mathbf{Prop}}
\newcommand{\Bool}{\mathbf{Bool}}

@@

\begin{example}
  The opposite category of every Kleene category is a model of a variant of
  conjunctive ordered logic, where the Kleene star plays the role of the ``of
  course'' modality from substructural logics.
\end{example}

@@

This result feels similar in spirit to the definition of action
algebras, which are algebras where the product also has adjoint
operations which results in the Kleene star being more easily
axiomatized \cite{kozen1994}. We are now ready to prove that our
concept of formal grammars fit nicely within our categorical
framework. We start by presenting a well-known construction
from presheaf categories.

@@

  By the lemma above, $\Set^{\cat{C}}$ is monoidal closed, and since it
  is a presheaf category, it has coproducts, and since the tensor
  is a left adjoint, it preserves colimits and, therefore, it is
  a distributive category.

@@

@@

To this end we define
\begin{enumerate}
\item $(\phi, S) \otimes (\psi, T) = (\phi \otimes \psi, S\otimes T)$ where
  \[ (S \otimes T)(w, p) = (w_1w_2 = w) \times (q_1,q_2 = p) \times S\,w_1\,q_1 \times T\,w_2\,q_2\]
\item $(\phi, S) \multimap (\psi, T) = (\phi \multimap \psi, S \multimap T)$ where
  \[ (S \multimap T)(w,p) = w' \to q \to S\,w'\,q \to T (ww') (p\,q) \]
\item $\mu X. ??$ ??
\end{enumerate}

>>>>

\newcommand{\cat}[1]{\mathbf{#1}}
\newcommand{\lto}{\multimap}
\newcommand{\Set}{\mathbf{Set}}
\newcommand{\Type}{\mathbf{Type}}
\newcommand{\Prop}{\mathbf{Prop}}
\newcommand{\Bool}{\mathbf{Bool}}

@@

\begin{example}
  The opposite category of every Kleene category is a model of a variant of
  conjunctive ordered logic, where the Kleene star plays the role of the ``of
  course'' modality from substructural logics which allows hypotheses to
  be discarded or duplicated.
\end{example}

@@

This result feels similar in spirit to the definition of action
algebras, which are algebras where the product also has adjoint
operations which results in the Kleene star being more easily
axiomatized \cite{kozen1994}.

We are now ready to prove that our concept of formal grammars fits
nicely within our categorical framework. We start by presenting a
well-known construction from presheaf categories.

@@

  By the lemma above, $\Set^{\cat{C}}$ is monoidal closed, and since it
  is a presheaf category, it has coproducts. Furthermore, the tensor
  is a left adjoint, i.e. it preserves colimits and, therefore, it is
  a distributive category.

@@

Much like in the posetal case, the abstract structure of a 2-Kleene
algebra is expressive enough to synthetically reason about formal
languages. A significant difference between them is that while Kleene
algebras can reason about language containment, 2-Kleene algebras can
reason about parsing as well.

For the rest of the paper we will use the presheaf category
$\Set^{\cat{\Sigma^*}}$ as a concrete model that will serve as our
guide when expanding abstract formalism presented in this section so
that it can handle more classes of languages beyond regular ones.

@@

To this end we define
\pedro{I'm pretty sure that these definitions have been somewhat presented
  in ``Glueing and orthogonality for models of linear logic'' - Hyland and Schalk.
We should probably cite them so that we don't have to rederive their constructions and proofs here.}
\begin{enumerate}
\item $(\phi, S) \otimes (\psi, T) = (\phi \otimes \psi, S\otimes T)$ where
  \[ (S \otimes T)(w, p) = (w_1w_2 = w) \times (q_1,q_2 = p) \times S\,w_1\,q_1 \times T\,w_2\,q_2\]
\item $(\phi, S) \multimap (\psi, T) = (\phi \multimap \psi, S \multimap T)$ where
  \[ (S \multimap T)(w,p) = w' \to q \to S\,w'\,q \to T (ww') (p\,q) \]
\item $\mu X. ??$ ??
\end{enumerate}
